A Survey of Near-Term Exoskeleton Technology\\ (Draft 4.0)

### 2015-10-01 control.tex
more updates

\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{cite}
\usepackage{graphicx}
% \usepackage[dvips]{graphicx}
% \usepackage{epsfig} % for postscript graphics files
  % \graphicspath{{../eps/}}
% \DeclareGraphicsExtensions{.eps}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage[cmex10]{amsmath}
%\usepackage{array}
%\usepackage{mdwmath}
%\usepackage{mdwtab}
%\usepackage{eqparbox}
\usepackage[tight,footnotesize]{subfigure}
%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
%\usepackage{fixltx2e}
%\usepackage{stfloats}
\usepackage{hyperref}

@@

It is useful to provide human programmers
with idealized abstractions of control of a single degree of
freedom. Here are some common alternatives that could be provided by low level control:
\begin{itemize}
\item
Position control. The joint moves to a specified position and stays there independent
of load.
\item
Velocity control. The joint moves at a specified velocity independent of load.
\item
Force or torque control. The joint provides a specified force or torque independent
of its motion.
\item
Impedance control: The user specifies a desired position (for a spring), and commands
a desired
stiffness, damping, and potentially a modified inertia or moment of inertia.
Modifying inertia is more dangerous than providing variable active stiffness
or damping.
Or (almost equivalently),
the user specifies a desired force or torque
which changes according to the position, velocity,
and acceleration of the joint.
\item
Valve control (for fluidic systems):
\begin{equation}
\vu = \mK_{\vq} (\vq - \vq_d) + \mK_{\dot{\vq}} ( \dot{\vq} - \dot{\vq}_d )
+ \mK_{\vtau} ( \vtau - \vtau_d )
+ \mK_{\dot{\vtau}} ( \dot{\vtau} - \dot{\vtau}_d ) + \vu_{\ff}
\end{equation}
where $\vu$ is a valve command vector (with a valve command for each actuator),
$\vq$ is the vector of joint angles, $\dot{\vq}$ are joint velocities,
$\vtau$ are actuator forces or joint torques, and $\dot{\vtau}$ is the rate of change
of the forces or torques (the time derivative of $\vtau$). 
The subscript $d$ indicates a desired value,
$\mK_{\vq}$ is a joint position gain matrix,
$\mK_{\dot{\vq}}$ is a joint velocity gain matrix, 
$\mK_{\vtau}$ is a joint force/torque gain matrix,
and $\mK_{\dot{\vtau}}$ is a joint force/torque derivative gain matrix.
$\vu_{\ff}$ are feedforward valve commands.
This formulation takes into account the most important hydraulic actuator dynamics,
and allows full state feedback
(including $\vtau$ as part of the state), pole placement, and linear quadratic
regulator (LQR) control of
a third order model of actuator plus rigid body dynamics.
If electric motor and power amplifier dynamics are taken into account in electric
motor systems, there would be an analogous equation for the commanded
motor voltage of
an analog power amplifier or
the control input of a PWM amplifier.
The torque error term $\mK_{\vtau} ( \vtau - \vtau_d )$ could be replaced or augmented
by a servo on piston differential pressure in a fluidic system or
an electrical current servo in an electric system.
\item
Online optimization of accelerations, actuator forces and torques, and contact forces
using quadratic programming. See Section \ref{sec:qp} for further explanation.
\end{itemize}

@@

\subsection{Behavior Selection}

@@

This section describes how to handle underactuation and overactuation using
``floating body dynamics.''
We have been sweeping this
issue under the rug so far. It is now useful to be more explicit.
There are two kinds of underactuation we want to consider. The first
is ``floating body dynamics'', where the robot is not bolted to the floor.
Contacts are made and broken during walking, running, and many other behaviors,
and contacts slip.
Humanoid robots (and humans) have floating body dynamics.
A ``root'' location (usually somewhere
on the pelvis) is chosen, and the position, velocity, orientation, and angular
velocity of that point and the rigid body it is on
are included in the state vector $\vq$. The corresponding dimensions of the
velocity vector $\dot{\vq}$ are not actuated. A Jacobian matrix
$\mJ_{\myx}$ is used to map from actuation variables $\vu$ to joint torques $\vtau$:
\begin{equation}
\vtau_{\myx} = \mJ_{\myx} \vu_{\myx}
\end{equation}
The rows in $\mJ_{\myx}$ corresponding to the root linear and angular velocities
are all zero.
To maintain compatibility in our joint representation, we will similarly
map operator ``muscle commands'' $\myo$ to operator joint torques $\vtau_{\myo}$.
\begin{equation}
\vtau_{\myo} = \mJ_{\myo} \vu_{\myo}
\end{equation}

@@

Integral admittance control is a refinement of 
sensitivity amplification control.
This section is quotes from:\\
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=7139989}

{\it [Exoskeletons] have the implicit property
of causing a virtual modification of the dynamic response of
the human limb. We use this property of the exoskeletons
action to formulate a unified control design framework called
Integral Admittance [torque to angle] Shaping, which designs exoskeleton 
controllers capable of producing the desired dynamic response
for the assisted limb. In this framework, a virtual increase
in the admittance of the limb is produced by coupling it
to an exoskeleton that exhibits active behavior. Specifically,
our framework shapes the magnitude profile of the integral
admittance (i.e. torque-to-angle relationship) of the coupled
human-exoskeleton system, such that the desired assistance is
achieved. This framework also ensures that the coupled stability
and passivity are guaranteed.}
%~\cite{Nagarajan_etal_2015}

{\it ... the impedance of the coupled human-exoskeleton
system needs to be reduced below that of the unassisted
human limb. This implies that the exoskeleton needs to
cancel its own impedance first and then compensate for
at least a part of the human limb’s impedance. Therefore,
the desired exoskeleton behavior must be that of a negative
impedance ... Consequently,
the feedback gains will all be
positive. In other words, the exoskeleton controller uses
positive feedback, and hence the exoskeleton exhibits active
behavior, which is capable of performing net positive work
on the limb. ...
However, positive feedback naturally raises the question of
stability, and so we now explain how coupled stability can be
achieved. Although the exoskeleton exhibits active behavior,
which can be potentially destabilizing, the controller can be
designed such that the coupled human-exoskeleton system
is stable and passive. ... 
}
%~\cite{Nagarajan_etal_2015}

{\it
... Inertia compensation is more complex ...
It can be shown that using only positive acceleration
feedback ..., the gain margin of the coupled system
reduces to the moment of inertia of the exoskeleton, which
implies that the exoskeleton controller ... can at the most
compensate for the exoskeleton’s own moment of inertia
before going unstable. This implies that the moment of
inertia of the coupled human-exoskeleton system cannot be
reduced below that of the unassisted human limb, without
compromising coupled stability. However, using low-pass
filtered acceleration feedback, it can be shown that inertia 
reduction can be achieved. ...
This work uses filtered acceleration
feedback with a second-order low-pass Butterworth filter. ...
}
%~\cite{Nagarajan_etal_2015}

@@

{\it The robot utilized the dual-mode control scheme, which is
comprised of the active control for the stance phase and the
passive control (using bypass valves) for the swing phase, to achieve high walking
speed in the swing phase while supporting heavy loads in
the stance phase. To reduce the sudden change of the torque
command at the transition from the swing phase to the stance
phase, a smoothing method is adopted. We also implemented
a pre-transition method to take a foot off quickly for fast
walking by predicting the change from the swing to the stance
in advance.}\\
%~\cite{IEEE07222598}
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=7222598}

@@

Stance: virtual joint torque control method Kazerooni et al.(2005)
When contact location between the wearer and the
exoskeleton is not fixed and difficult to estimate, this method
has been shown to be an effective method to generate the
locomotion for an exoskeleton robot.[3],[4],[21]
\begin{verbatim}
H. Kazerooni , Z. Racine , L. Huang and R. Steger ”On the control of
the Berkeley lower extremity exoskeleton (BLEEX),” Proc. IEEE Int.
Conf. Robot. Autom., pp. 4364-4371, 2005.
Racine Jean-Louis Charles ”Control of a Lower Extrmity Exoskeleton
for Human Performance Amplification,” University of California,2003.
582
Xiuxia Yang, Hongchao Zhao, Yi Zhang, Xiaowei Liu ”Carrying
Lower Extreme Exoskeleton Rapid Terminal Sliding-Mode Robust
Control,” Journal of computers, vol.9, No.1,202-208. 2014.
\end{verbatim}

To reduce
sudden changes(command jump) at the phase transitions, a
smoothing method [5],[19] is introduced and a pre-transition
method is used to solve the swing delay due to the internal
pressure(approx 5 bar).
\begin{verbatim}
H. Kazerooni , Ryan Steger, Lihua Huang ”Hybrid Control of the
Berkeley Lower Extremity Exoskeleton (BLEEX),” Int. Journal of
Robotics Research, pp. 561-573, 2006.
oonbum Bae, Kyoungchul Kong, Masayoshi Tomizuka ”Gait Phase-
Based Control for a Rotary Series Elastic Actuator Assisting the knee
\end{verbatim}

>>>>

\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{cite}
\usepackage{graphicx}
% \usepackage[dvips]{graphicx}
% \usepackage{epsfig} % for postscript graphics files
  % \graphicspath{{../eps/}}
% \DeclareGraphicsExtensions{.eps}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage[cmex10]{amsmath}
%\usepackage{array}
%\usepackage{mdwmath}
%\usepackage{mdwtab}
%\usepackage{eqparbox}
\usepackage[tight,footnotesize]{subfigure}
%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
%\usepackage{fixltx2e}
%\usepackage{stfloats}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
\usepackage{hyperref}

@@

It is useful to provide human programmers
with idealized abstractions of control of a single degree of
freedom. Here are some common alternatives that could be provided by low level control:
\begin{itemize}
\item
Position control. The joint moves to a specified position and stays there independent
of load.
\item
Velocity control. The joint moves at a specified velocity independent of load.
\item
Force or torque control. The joint provides a specified force or torque independent
of its motion.
\item
Impedance control: The user specifies a desired position (for a spring), and commands
a desired
stiffness, damping, and potentially a modified inertia or moment of inertia.
Modifying inertia is more dangerous than providing variable active stiffness
or damping.
Or (almost equivalently),
the user specifies a desired force or torque
which changes according to the position, velocity,
and acceleration of the joint.
\item
Valve control (for fluidic systems)
\begin{equation}
\vu = \mK_{\vq} (\vq - \vq_d) + \mK_{\dot{\vq}} ( \dot{\vq} - \dot{\vq}_d )
+ \mK_{\vtau} ( \vtau - \vtau_d )
+ \mK_{\dot{\vtau}} ( \dot{\vtau} - \dot{\vtau}_d ) + \vu_{\ff}
\label{eq:valve}
\end{equation}
where $\vu$ is a valve command vector (with a valve command for each actuator),
$\vq$ is the vector of joint angles, $\dot{\vq}$ are joint velocities,
$\vtau$ are actuator forces or joint torques, and $\dot{\vtau}$ is the rate of change
of the forces or torques (the time derivative of $\vtau$). 
The subscript $d$ indicates a desired value,
$\mK_{\vq}$ is a joint position gain matrix,
$\mK_{\dot{\vq}}$ is a joint velocity gain matrix, 
$\mK_{\vtau}$ is a joint force/torque gain matrix,
and $\mK_{\dot{\vtau}}$ is a joint force/torque derivative gain matrix.
$\vu_{\ff}$ are feedforward valve commands.
This formulation takes into account the most important hydraulic actuator dynamics,
and allows full state feedback
(including $\vtau$ as part of the state), pole placement, and linear quadratic
regulator (LQR) control of
a third order model of actuator plus rigid body dynamics.

If electric motor and power amplifier dynamics are taken into account in electric
motor systems, there would be an analogous equation for the commanded
motor voltage of
an analog power amplifier or
the control input of a PWM amplifier.
The torque error term $\mK_{\vtau} ( \vtau - \vtau_d )$ could be replaced or augmented
by a servo on piston differential pressure in a fluidic system or
an electrical current servo in an electric system.

Often this type of control is referred to as a ``torque control inner loop''
with a ``position/velocity control outer loop''. In this approach
the inner and outer loops are combined in one equation (eq.~\ref{eq:valve}),
and there is no
longer ``inner'' and ``outer'' control quantities. All feedback is
treated the same way.
\item
Online optimization of accelerations, actuator forces and torques, and contact forces
using quadratic programming. See Section \ref{sec:qp} for further explanation.
\end{itemize}

@@

\subsection{``Sentence'' Control: Behavior Selection}

@@

This section describes how to handle underactuation and overactuation using
``floating body dynamics.''
We have been sweeping this
issue under the rug so far. It is now useful to be more explicit.
There are two kinds of underactuation we want to consider. The first
is ``floating body dynamics'', where the robot is not bolted to the floor.
Contacts are made and broken during walking, running, and many other behaviors,
and contacts slip.
Humanoid robots (and humans) have floating body dynamics.
A ``root'' location (usually somewhere
on the pelvis) is chosen, and the position, velocity, orientation, and angular
velocity of that point and the rigid body it is on
are included in the state vector $\vq$. The corresponding dimensions of the
velocity vector $\dot{\vq}$ are not actuated. A Jacobian matrix
$\mJ_{\myx}$ is used to map from actuation variables $\vu$ to joint torques $\vtau$:
\begin{equation}
\vtau_{\myx} = \mJ_{\myx} \vu_{\myx}
\end{equation}
The rows in $\mJ_{\myx}$ corresponding to the root linear and angular velocities
are all zero.
To maintain compatibility in our joint representation, we will similarly
map operator ``muscle commands'' $\vu_{\myo}$ to operator joint torques $\vtau_{\myo}$.
\begin{equation}
\vtau_{\myo} = \mJ_{\myo} \vu_{\myo}
\end{equation}

@@

Integral admittance control is a refinement of 
sensitivity amplification control.
This section quotes from:\\
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=7139989}

{\it ``[Exoskeletons] have the implicit property
of causing a virtual modification of the dynamic response of
the human limb. We use this property of the exoskeletons
action to formulate a unified control design framework called
Integral Admittance [torque to angle] Shaping, which designs exoskeleton 
controllers capable of producing the desired dynamic response
for the assisted limb. In this framework, a virtual increase
in the admittance of the limb is produced by coupling it
to an exoskeleton that exhibits active behavior. Specifically,
our framework shapes the magnitude profile of the integral
admittance (i.e. torque-to-angle relationship) of the coupled
human-exoskeleton system, such that the desired assistance is
achieved. This framework also ensures that the coupled stability
and passivity are guaranteed.}
%~\cite{Nagarajan_etal_2015}

{\it ``... the impedance of the coupled human-exoskeleton
system needs to be reduced below that of the unassisted
human limb. This implies that the exoskeleton needs to
cancel its own impedance first and then compensate for
at least a part of the human limb’s impedance. Therefore,
the desired exoskeleton behavior must be that of a negative
impedance ... Consequently,
the feedback gains will all be
positive. In other words, the exoskeleton controller uses
positive feedback, and hence the exoskeleton exhibits active
behavior, which is capable of performing net positive work
on the limb. ...
However, positive feedback naturally raises the question of
stability, and so we now explain how coupled stability can be
achieved. Although the exoskeleton exhibits active behavior,
which can be potentially destabilizing, the controller can be
designed such that the coupled human-exoskeleton system
is stable and passive. ... 
}
%~\cite{Nagarajan_etal_2015}

{\it
``... Inertia compensation is more complex ...
It can be shown that using only positive acceleration
feedback ..., the gain margin of the coupled system
reduces to the moment of inertia of the exoskeleton, which
implies that the exoskeleton controller ... can at the most
compensate for the exoskeleton’s own moment of inertia
before going unstable. This implies that the moment of
inertia of the coupled human-exoskeleton system cannot be
reduced below that of the unassisted human limb, without
compromising coupled stability. However, using low-pass
filtered acceleration feedback, it can be shown that inertia 
reduction can be achieved. ...
This work uses filtered acceleration
feedback with a second-order low-pass Butterworth filter. ...''
}
%~\cite{Nagarajan_etal_2015}

@@

{\it ``The robot utilized the dual-mode control scheme, which is
comprised of the active control for the stance phase and the
passive control (using bypass valves) for the swing phase, to achieve high walking
speed in the swing phase while supporting heavy loads in
the stance phase. To reduce the sudden change of the torque
command at the transition from the swing phase to the stance
phase, a smoothing method is adopted. We also implemented
a pre-transition method to take a foot off quickly for fast
walking by predicting the change from the swing to the stance
in advance.''}\\
%~\cite{IEEE07222598}
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=7222598}

@@

Stance: virtual joint torque control method Kazerooni et al.(2005)
When contact location between the wearer and the
exoskeleton is not fixed and difficult to estimate, this method
has been shown to be an effective method to generate the
locomotion for an exoskeleton robot.
\begin{verbatim}
Xiuxia Yang, Hongchao Zhao, Yi Zhang, Xiaowei Liu ”Carrying
Lower Extreme Exoskeleton Rapid Terminal Sliding-Mode Robust
Control,” Journal of computers, vol.9, No.1,202-208. 2014.
\end{verbatim}

To reduce
sudden changes(command jump) at the phase transitions, a
smoothing method is introduced and a pre-transition
method is used to solve the swing delay due to the internal
pressure(approx 5 bar).
\begin{verbatim}
oonbum Bae, Kyoungchul Kong, Masayoshi Tomizuka ”Gait Phase-
Based Control for a Rotary Series Elastic Actuator Assisting the knee
\end{verbatim}

### 2015-10-01 intent.tex
more updates

\begin{enumerate}
\item
We recommend using a Bayesian {\bf (probabilistic) framework}
for intent recognition and prediction.
\item
We recommend developing an intent prediction system that can use
as {\bf many sensors and measurements} of the operator as are possible.
\item
We recommend each exoskeleton controller
to be used by and {\bf optimized for a single operator.}
A substantial investment in capturing the operators normal behavior,
operator training and learning, and controller customization should be made.
\end{enumerate}

@@

@@

@@

Detection reaction time (the delay between an anticipated 
stimulus and the onset of movement) 
for the operators we are interested in (young adults)
is about
190ms for visual input and
160ms for auditory stimulus in a detection task.
It takes longer to recognize a signal from a set of ongoing events,
and even longer if the operator has to make a choice based on the signal
(go left or right, for example).
Reaction times for complex actions are longer that reaction times
for simple actions.\\
\url{https://en.wikipedia.org/wiki/Mental_chronometry}.

@@

The second view is that the field is full of crap. In the above demo,
the user looked at LEDs in order to get a measurable brain signal.
This means that one could have just tracked where the user was looking
(no brain signals used at all) to figure out what to do.
In general the bit rate coming out of any brain-machine interface is
very very low, so decoding the signal is going to be very very slow
(probably longer than the reaction time of an athletic subject).

Nevertheless, a middle of the road view might be beneficial. It is possible
that brain signals,
in the context of a wide range of other measurements,
might improve overall intent estimation.
It may be that brain signals are useful to tell us {\bf when} a behavior
is happening, if they are not very useful for telling us {\bf what}
behavior is happening.

@@

Electromyography 
signals are noisy and it is not easy to extract
meaningful data from the raw signal. 
EMG sensors
need to be attached to the operator, calibrated,
and periodically recalibrated during use, which takes time.
Perspiration changes the measurement. 
Muscle movement leads to surface electrodes losing good contact,
fully detaching, or
even ripping off (especially for large fast
movements, the worst time).

Off the shelf electronics:\\
\url{http://www.advancertechnologies.com/p/muscle-sensor-v3.html}

@@

@@

In a full-body exoskeleton
(not necessarily fully actuated), the
exoskeleton can be used to measure positions, velocities,
and accelerations. For partial exoskeletons,
it may be the case that additional sensors such as distributed 
angle sensors, tachometers, or inertial measurement units (IMUs)
are attached to the operator or part of the operator's clothing.

@@

We note that the operator will use more command channels that just
body movement, since
the operator controls more than the pose and motion of the
exoskeleton. What if there are ``pointing'' sensors or communication
devices that need to be aimed at or track an area of interest? What if
there are additional (physical or virtual) pan/tilt/zoom cameras
pointing to the side and rear or line of sight secure communications,
for example. What about other controls, such as power assist level or
thermal control? It is an open question
how an operator can naturally express intent and
control these additional degrees of freedom.

@@

3) We recommend each exoskeleton controller
to be used by and optimized for a single operator.
A substantial investment in capturing the operators normal behavior,
operator training and learning, and controller customization should be made.

>>>>

\begin{enumerate}
\item
We recommend using a Bayesian {\bf (probabilistic) framework}
for intent recognition and prediction.
\item
We recommend developing an intent prediction system that can use
as {\bf many sensors and measurements} of the operator as possible.
\item
We seriously consider {\bf implanting devices and markers} in operators.
\item
We recommend each exoskeleton controller
to be used by and {\bf optimized for a single operator.}
A substantial investment in capturing the operator's normal behavior,
operator training and learning, and controller customization should be made.
\end{enumerate}

@@

\subsection{Operator state (joint positions and velocities)}

@@

Estimating operator physical state: how much fatigue (perhaps down
to the level of metabolic and muscle biochemistry estimates) and
current and future capabilities.

@@

Detection reaction time (the delay between an anticipated 
stimulus and the onset of movement) 
for the operators we are interested in (young adults)
is about
190ms for visual input and
160ms for auditory stimulus in a detection task.
It takes longer to recognize a signal from a set of ongoing events,
and even longer if the operator has to make a choice based on the signal
(go left or right, for example).
Reaction times for complex actions are longer than reaction times
for simple actions.\\
\url{https://en.wikipedia.org/wiki/Mental_chronometry}.

@@

The second view is that the field is full of crap. In the above demo,
the user looked at LEDs in order to get a measurable brain signal.
This means that one could have just tracked where the user's eyes were looking
(no brain signals used at all) to figure out what to do.
In general the bit rate coming out of any brain-machine interface is
very very low, so decoding the signal is going to be very very slow
(probably longer than the reaction time of an athletic subject).

Nevertheless, a middle of the road view might be beneficial. It is possible
that brain signals,
in the context of a wide range of other measurements,
might improve overall intent estimation.
It may be that brain signals are useful to tell us {\bf when} a behavior
is going to happen, if they are not very useful for telling us {\bf what}
behavior is going to happen.

@@

Electromyography 
signals are noisy and it is not easy to extract
meaningful data from the raw signal. 
EMG sensors
need to be attached to the operator, calibrated,
and periodically recalibrated during use, which takes time.
Perspiration changes the measurement. 
Muscle movement leads to surface electrodes losing good contact,
fully detaching, or
even ripping off (especially for large fast
movements, the worst time). Muscle moves relative to skin,
so any electrodes going through the skin will get ripped off
in large fast movements.

Implanted electrodes
with wireless connections for signals and power
get high quality stable measurements.
For the class of users whose lives depend on the exoskeleton,
it might be worth implanting devices, if they did not interfere
with performance (soreness, pain in some positions or due to
movement). It is roughly equivalent to
operating with embedded shrapnel that has healed over.
A research goal would be getting these electrodes the size of rice
grains or smaller.

Implantable EMG electrodes (Figure~\ref{f:emg}):\\
\url{http://neural.iit.edu/research/imes/}\\
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=4716669}\\
\url{http://link.springer.com/chapter/10.1007\%2F978-3-319-08072-7_12}\\
\url{http://www.ncbi.nlm.nih.gov/pubmed/19224729}\\
\url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2878855/}\\

\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{figs/emg.png}
\caption{Implanted EMG electrodes \url{http://neural.iit.edu/research/imes/}}
\label{f:emg}
\end{figure}

Off the shelf electronics for surface EMG:\\
\url{http://www.advancertechnologies.com/p/muscle-sensor-v3.html}

\subsection{Operator muscle motion}

This implantable device measures relative motion of muscle relative
to other muscles and soft tissues:\\
\url{http://bme.usc.edu/assets/004/54640.pdf}

It would be relatively easy to develop implanted MEMs IMU devices (MEMs accelerometers
and gyros) to measure high frequency events (movement onset, reflexes,
impacts, relative muscle moition, muscle contraction, ...).\\
\url{http://www.designworldonline.com/surface-mount-accelerometer-for-implantable-medical-devices/}\\
\url{https://www.endevco.com/meggitt-sensing-systems-announces-endevco\%C2\%AE-surface-mount-accelerometer-series-for-implantable-medical-device-applications/}\\
\url{http://www.sciencedirect.com/science/article/pii/S0924424713005979}\\
\url{http://www.ncbi.nlm.nih.gov/pubmed/17574533}\\
\url{http://circep.ahajournals.org/content/3/6/657.full}\\

@@

These devices measure relative motion of tissue, and could be used
to measure force by measuring tendon stretch:\\
\url{http://bme.usc.edu/assets/004/54640.pdf}

@@

In a full-body exoskeleton
(not necessarily fully actuated), the
exoskeleton can be used to measure positions, velocities,
and accelerations. For partial exoskeletons,
it may be the case that additional sensors such as distributed 
angle sensors, tachometers, or inertial measurement units (IMUs)
are attached to or implanted in the operator 
or part of the operator's clothing.

@@

We note that the operator will probably use more command channels that just
body movement, since
the operator controls more than the pose and motion of the
exoskeleton. What if there are ``pointing'' sensors or communication
devices that need to be aimed at or track an area of interest? What if
there are additional (physical or virtual) pan/tilt/zoom cameras
pointing to the side and rear or line of sight secure communications,
for example. What about other controls, such as power assist level or
thermal control? It is an open question
how an operator can naturally express intent and
control these additional degrees of freedom.

@@

3) We seriously consider implanting devices and markers in operators.

4) We recommend each exoskeleton controller
to be used by and optimized for a single operator.
A substantial investment in capturing the operators normal behavior,
operator training and learning, and controller customization should be made.

### 2015-10-01 technology.tex
more updates

Another useful specification is the cost in torque output and power
of exoskeleton and payload mass if the exoskeleton runs like a human.
Note that the speed requirements on the exoskeleton do not change as
payload increases.
\begin{itemize}
\item
hip: 2Nm/kg, 13W/kg
\item
knee 1.5Nm/kg, knee 6W/kg
\item
ankle 2Nm/kg, 13W/kg
\end{itemize}
From: {\it The biomechanics of running,} Tom F. Novacheck,
Gait and Posture 7 (1998) 77–95.

@@

However, on the Atlas robot differential oil pressure is used
to estimate actuator force, as this form of force sensing
is quite mechanically robust (low maintenance).
This measurement is before the piston
seals and transmission, so it does not capture these sources
of friction. The joint torques are off by up to 10Nm at each joint.
The Sarcos robot uses load cells in the actuator to joint transmissions.

On the Atlas humanoid robot there was electromagnetic interactions between
nearby valves.

@@

The JST lab at the University of Tokyo pushed the design of water
cooled electric motors and motor drivers. This led to the high
performance of the Schaft robot, a spinoff from that lab.

The water cooled motor driver they developed is capable of 80V and
80A continuous.\\
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=5649683}\\

They modified
MAXON EC-powermax30 200W brush-less motors, and added harmonic drives
to make an integrated actuator.
The maximum speed of these actuators was 3 revolutions/second.
The maximum torque using 2 drivers/motor was theoretically 400Nm, but
we have only seen 300Nm in the published data. 2 rev/s can be achieved
at peak torque (Fig~\ref{f:s1}).\\
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=5649683}\\

@@

@@

This effort ``went dark'' after Google bought the spinoff company
Schaft, so there is a lot we don't know.

@@

Different types of actuation can be combined: springs and electric motors,
for example. This combination can also be viewed as combining energy
storage and actuation. Slow pneumatics and fast electric motors.\\
\url{http://www.cs.cmu.edu/~cga/exo/survey.pdf}\\
Mini/Macro.\\
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=4479941}\\
In this view, the human operator is just
one more set of actuators to be added to the combined system.

@@

{\bf Do we need physical contact?}
Current exoskeletons use physical contact with the operator to control
the exoskeleton. Can we make an exoskeleton that does not touch or
minimally touches the operator, and in which there is no power
transfer between the operator and the exoskeleton? These types of
systems might allow an operator to behave more normally, have a
greater range of movement, be less tiring, and for non-enclosing
systems, allow the operator to move independently of the system (such
as aim and fire standard weapons or dive for cover) for additional
operator performance and safety.
\begin{enumerate}
\item
A "fat suit" or "sumo suit" could fully enclose the operator, but
use distance sensing to move with the operator.
\item
A "shield" could enclose the front of an operator (or whatever
portion of the operator is threatened) and move with the operator,
\item
A "human shield" (actually a machine shield) could move ahead of
an operator, mimicking the operators movement.
\item
An "angel" could walk, jump, or fly between an operator and potential
hostile sites, and deflect or disable projectiles.
\end{enumerate}

>>>>

Another useful specification is the cost in torque output and power
of exoskeleton and payload mass if the exoskeleton runs like a human.
Note that the speed requirements on the exoskeleton do not change as
payload increases.
\begin{itemize}
\item
hip: 2Nm/kg, 13W/kg
\item
knee 1.5Nm/kg, knee 6W/kg
\item
ankle 2Nm/kg, 13W/kg
\end{itemize}
From: {\it The biomechanics of running,} Tom F. Novacheck,
Gait and Posture 7 (1998) 77–95.
\url{http://www.cs.cmu.edu/~cga/exo/info/Novacheck1998.pdf}

@@

However, on the Atlas robot differential oil pressure is used
to estimate actuator force, as this form of force sensing
is quite mechanically robust (low maintenance).
This measurement is before the piston
seals and transmission, so it does not capture these sources
of friction. The joint torques are off by up to 10Nm at each joint.
The Sarcos robot uses load cells in the actuator to joint transmissions,
and gets much better measurements. In 10 years we have only had
one load cell (of 30) fail. It was undersized due to a design error.

On the Atlas humanoid robot there was electromagnetic interactions between
nearby valves which required magnetic shielding.

@@

The JST lab at the University of Tokyo pushed the design of water
cooled electric motors and motor drivers. This led to the high
performance of the Schaft robot, a spinoff from that lab.
We note that the team that won the DARPA Robotics Challenge, KAIST, used
forced air cooling of its electric motors and drive circuits,
rather than using water cooling. More details follow.

The water cooled motor driver JST developed is capable of 80V and
80A continuous.\\
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=5649683}\\

JST modified
MAXON EC-powermax30 200W brush-less motors, and added harmonic drives
to make an integrated actuator.
The maximum speed of these actuators was 3 revolutions/second.
The maximum torque using 2 drivers/motor was theoretically 400Nm, but
we have only seen 300Nm in the published data. 2 rev/s can be achieved
at peak torque (Fig~\ref{f:s1}).\\
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=5649683}\\

@@

Here is information on the KAIST robot DRC-HUBO with air cooled electric
motors:

{\it ``With customized motor drivers and an air cooling system (fans and fins), the robot is able to drive 3x to 4x more current than what is listed in the motor specs, with peaks of 30 amperes in some cases.}

{\it ``Compliance: The team wanted to make their robot compliant, but didn't want to use force-torque sensors and a conventional feedback controller (which they feared would introduce instabilities). So instead the implemented compliance on their custom motor driver, using a special amplifier.}

{\it ``When motors draw lots of current, the main power system may fail to provide enough power to critical components. To avoid that, the team used a supercapacitor system that keeps computers, communications, and some sensors like gyros running even if the main power system goes down.''}\\
\url{http://spectrum.ieee.org/automaton/robotics/humanoids/how-kaist-drc-hubo-won-darpa-robotics-challenge}

@@

The JST effort ``went dark'' after Google bought the spinoff company
Schaft, so there is a lot we don't know.

We will get more information about the DRC-HUBO robot in November at
the Humanoids conference.

@@

Different types of actuation can be combined: springs and electric motors,
for example. This combination can also be viewed as combining energy
storage and actuation. XoR combined slow pneumatics and fast electric motors.\\
\url{http://www.cs.cmu.edu/~cga/exo/survey.pdf}\\
Mini/Macro.\\
\url{http://ieeexplore.ieee.org/xpl/login.jsp?tp=&arnumber=4479941}\\
In this view, the human operator is just
one more set of actuators to be added to the combined system.

@@

{\bf Do we need physical contact?}
Current exoskeletons use physical contact with the operator to control
the exoskeleton. Can we make an exoskeleton that does not touch or
minimally touches the operator, and in which there is no power
transfer between the operator and the exoskeleton? These types of
systems might allow an operator to behave more normally, have a
greater range of movement, be less tiring, and for non-enclosing
systems, allow the operator to move independently of the system (such
as aim and fire standard weapons or dive for cover) for additional
operator performance and safety.
\begin{enumerate}
\item
A "fat suit" or "sumo suit" could fully enclose the operator, but
use distance sensing to move with the operator.
\item
A "shield" could enclose the front of an operator (or whatever
portion of the operator is threatened) and move with the operator,
\item
A "human shield" (actually a machine shield) could move ahead of
an operator, mimicking the operators movement.
\item
An "angel" could walk, jump, or fly between an operator and potential
hostile sites, and deflect or disable projectiles.
\end{enumerate}
A variant of this approach is to have an exoskeleton that is attached
to the operator with springs, or springs and dampers. This could give
the system much more flexibility in terms of deviation between
operator and exoskeleton motion, and provide a robust substrate for
operator-exoskeleton force sensing. Maybe a tight fitting exoskeleton
is a bad idea.
